% MeshCore for Bitcoiners — a Bitcoin Austria talk.
% Build from the repo ROOT:  make           (or: latexmk -xelatex meshcore-for-bitcoiners.tex)
% Requires the `theme/` submodule (Bitcoin Austria beamer style). If it is
% missing, run:  git submodule update --init --recursive
%
% NOTE: this is a working draft built from handoff-20260626.md; develop further.
\documentclass[aspectratio=169,t,9pt]{beamer}

\usepackage{theme/bitcoin-austria}

% Talk-specific helper: a Bitcoin vs MeshCore comparison with a "where the
% analogy breaks" callout (wraps the theme's generic \comparisonslide).
% Pass an empty 4th argument {} to omit the callout.
\newcommand{\bridgeslide}[4]{%
  \comparisonslide{#1}{Bitcoin}{#2}{MeshCore}{#3}{Where it breaks}{#4}}

\title{MeshCore for Bitcoiners}
\subtitle{Permissionless communication — from money to radio}
\author{Harald Schilly — Bitcoin Austria}
\date{\today}

\begin{document}

\begin{frame}[plain,noframenumbering]
  \titlepage
\end{frame}

% ============================================================ what is MeshCore?
% Bitcoiners don't know MeshCore — introduce it before drawing the parallels.
\fillerslide[A radio mesh you can join without asking]{What is MeshCore?}

\begin{frame}{Messaging without the network}
  \begin{itemize}
    \item Off-grid text messaging over long-range \textbf{LoRa} radio —
      no internet, no cell tower, no SIM
    \item Open-source firmware running on cheap LoRa boards
    \item Uses \textbf{licence-free} ISM radio bands — worldwide, region by
      region (EU 868\,MHz; US/AU 915\,MHz; \dots)
  \end{itemize}
  \vfill
  \begin{alertblock}{The EU angle}
    On these bands \emph{anyone} may transmit — no licence, no registration, no
    fee. You only respect the power and duty-cycle limits.
  \end{alertblock}
\end{frame}

\begin{frame}{Where it comes from}
  \begin{itemize}
    \item Grew out of the \textbf{Meshtastic} world — then rebuilt from scratch
    \item Created by \textbf{Scott Powell} in \textbf{late 2024};
      web/mobile clients by Liam Cottle
    \item Andy Kirby — early tester and promoter (YouTube)
    \item A separate network — \emph{not} interoperable with Meshtastic
    \item Project home: \textbf{meshcore.io}
  \end{itemize}
\end{frame}

\begin{frame}{The hardware — tens of euros, not hundreds}
  \begin{itemize}
    \item \textbf{Companion} (your device: screen + Bluetooth) — SenseCAP
      \textbf{T1000-E}, credit-card size with GPS, $\sim$\,40\,EUR
    \item \textbf{Portable with a display} — Heltec LoRa\,32 V3 (OLED,
      $\sim$\,20\,EUR), Seeed Wio Tracker L1 Pro (screen, GPS, case)
    \item \textbf{Repeater} — SenseCAP Solar Node \textbf{P1-Pro}: 5\,W solar +
      battery, runs for days off-grid, pole-mounted; antenna + device
      $\sim$\,130\,EUR
    \item \textbf{Build your own} — Heltec / RAK WisBlock boards
  \end{itemize}
\end{frame}

% MeshCore (red, the subject) vs Meshtastic — generic comparison, not \bridgeslide
\comparisonslide{MeshCore vs Meshtastic}%
  {MeshCore}{%
    \begin{itemize}
      \item Role-based: companion / repeater / room server
      \item \textbf{Source-routed} along discovered paths
      \item Quieter, more scalable mesh
    \end{itemize}}%
  {Meshtastic}{%
    \begin{itemize}
      \item Easy to join, phone-first
      \item Flood-based — every node rebroadcasts
      \item Gets noisy as it grows
    \end{itemize}}%
  {Same cheap hardware}{%
    Same boards, same chips — a different firmware and routing philosophy.}

% A real packet route traced across Austria, captured from the observer.
\fullbleedslide[A packet routed across Austria \;\textbullet\; logger-at.meshcore.observer]{observer-package-across-austria-20260626}

% ---------------------------------------------------------------- core thesis
\begin{frame}{The core thesis}
  \begin{itemize}
    \item A self-generated \textbf{keypair} replaces institutional identity.
    \item \textbf{Bitcoin} replaces the bank account — gatekeeper: the bank.
    \item \textbf{MeshCore} replaces the SIM card / call sign — gatekeepers:
      the telco \emph{and} the state.
    \item No registration, no permission: you simply become a full participant.
  \end{itemize}
  \vfill
  \begin{alertblock}{The line the talk hangs on}
    Radio spectrum is heavily licensed. A permissionless, self-issued identity
    broadcasting in the ether is arguably even more subversive than
    permissionless money.
  \end{alertblock}
\end{frame}

% ---------------------------------------------------------------- identity
\fillerslide[The strongest parallel]{Identity is a keypair}

\bridgeslide{Identity is a keypair}{%
  \begin{itemize}
    \item Self-generated keypair, no issuer
    \item Address = hash of the public key
    \item ``Not your keys, not your coins''
    \item Verify the address out-of-band
  \end{itemize}}{%
  \begin{itemize}
    \item Self-generated keypair, on-device
    \item Routing target = public-key prefix
    \item ``Not your keys, not your identity''
    \item Verify keys via QR / word list (TOFU)
  \end{itemize}}{%
  HD wallets rotate addresses for privacy; a MeshCore identity is deliberately
  \emph{sticky} — you want to stay discoverable. No HD equivalent.}

% ---------------------------------------------------------------- signing
\fillerslide[The elegant twist]{The signing model}

\bridgeslide{Opposite jobs for the same keypair}{%
  \begin{itemize}
    \item Content is \textbf{public} (on-chain)
    \item The key protects \textbf{authorization}
    \item No valid signature → the transaction does not exist
  \end{itemize}}{%
  \begin{itemize}
    \item Identity is \textbf{public} (the advert)
    \item The key protects \textbf{content}
    \item Adverts are signed (Ed25519); DMs are encrypted (X25519)
  \end{itemize}}{%
  Group / channel messages use a pre-shared key — no per-sender signature.
  Anyone with the PSK can read \emph{and} write. (An honest flaw to name.)}

% ---------------------------------------------------------------- throughput
\fillerslide[A chosen constraint, not a weakness]{Throughput}

% TODO (verify before stage): exact EU 868 duty-cycle % and sub-band (~869 MHz).
\bridgeslide{The throttle is imposed, not a failure}{%
  \begin{itemize}
    \item Block size is a \textbf{consensus} cap
  \end{itemize}}{%
  \begin{itemize}
    \item The LoRa duty cycle is a \textbf{regulatory} cap (EU 868\,MHz band)
  \end{itemize}}{%
  In both cases the throttle is imposed, not ``they couldn't do more'' — and a
  deliberate cap is exactly what forces a second layer to exist.}

% ---------------------------------------------------------------- lightning
\fillerslide[Two distinct analogies — keep them separate]{Lightning Network}

\bridgeslide{Angle A — route along a discovered path}{%
  \begin{itemize}
    \item Find a route through specific channels
    \item Don't broadcast the payment to everyone
  \end{itemize}}{%
  \begin{itemize}
    \item An advert floods once; path info accumulates
    \item Later DMs are source-routed along that path
  \end{itemize}}{%
  Lightning onion-routes (hop privacy) and carries atomic value (HTLCs);
  MeshCore path bytes are in the clear and carry data — no settlement.}

\begin{frame}{Angle B — bootstrap, then jump layers}
  \begin{itemize}
    \item A small mesh message = the rare, expensive anchoring event
    \item Exchange a key / rendezvous over the mesh
    \item Move the bulk transfer to a fat pipe (internet, CB radio) — off-mesh
  \end{itemize}
  \vfill
  \begin{alertblock}{Where it breaks}
    Lightning stays cryptographically chained to L1 (you can unilaterally
    close). The mesh\,→\,fat-pipe handoff is a \textbf{one-time bootstrap}, not
    a trust-anchored channel — closer to WebRTC signaling or a BitTorrent
    tracker.
  \end{alertblock}
\end{frame}

% ---------------------------------------------------------------- facts
\fillerslide[Use these — they're checked]{Verified facts}

\begin{frame}{Crypto, checked}
  \begin{itemize}
    \item \textbf{Curves:} Ed25519 (signing) + X25519 (ECDH) — \emph{not} secp256k1
    \item \textbf{Identity:} 32-byte private + 32-byte public key, persistent
    \item \textbf{Routing prefix:} 1 byte originally (254 IDs → collisions);
      1–3 bytes since firmware 1.14+
    \item \textbf{Encryption:} group channels use AES-256-CTR with a PSK; the
      public channel uses a well-known weak key (convenience, not security)
  \end{itemize}
\end{frame}

\fillerslide[Questions?]{Thank you}

\end{document}
